A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2020-03-18 paper/paper - revision.Rnw
Flesh out the new outline

\begin{abstract}
Currently, to assess similarity in forensic shoeprint analysis, examiners must manually identify and label features in images of shoe prints or outsoles. In this paper, we discuss an algorithm for automatically labeling shoe outsole images using the same feature space examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals. The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes. Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast. This new approach represents a significant improvement to the current manual classification of footwear tread patterns, and could be integrated into current matching databases to reduce the amount of time examiners spend on data entry.
\end{abstract}

@@

In the US, real forensic analysis of shoe print evidence is varied and difficult. It is often impossible to conduct an individualizing analysis when a print is partial or low quality. Beyond the question of identification, it is more common (and arguably more tractable) to consider class characteristics. At the class level,  one question of interest \mt{motivate this more?} is calculating random match probability: Given a questioned shoe print, how many shoes in the relevant population have a similar tread pattern? We consider this question as distinct from make and model determination, since many shoes of different makes and models can still have very similar tread patterns (e.g., knockoffs, same models over different years, etc)

From a statistical perspective, random match probability could be calculated by sampling many shoes from a given population and counting the number of shoes in the sample that are "similar" to the given print. In practice, one potential method of sampling would be a low-profile scanner placed in a high traffic area. Such a scanner could take pictures of shoes walking past, and those images could then be used to create a database of shoes relevant to the local area. \mt{Are cameras more robust/weatherproof than pressure scanners? Why images instead of EverOS impressions?} \mt{What is the best way to say that such a scanner already exists, and/or that the grant is approved?} Assuming that such a database of images is available, the task becomes assessing the similarity of prints. This is done by identifying relevant features in the scans and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.

>>>>

\begin{abstract}
\mt{Update framing}
Currently, to assess similarity in forensic shoeprint analysis, examiners must manually identify and label features in images of shoe prints or outsoles. In this paper, we discuss an algorithm for automatically labeling shoe outsole images using the same feature space examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals. The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes. Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast. This new approach represents a significant improvement to the current manual classification of footwear tread patterns, and could be integrated into current matching databases to reduce the amount of time examiners spend on data entry.
\end{abstract}

@@

In the United States, forensic analysis of shoe print evidence in criminal cases is varied and difficult. The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis when a print is partial or of low quality. In cases where individualizing characteristics are not considered (estimated at 95\% of cases in the United States according to some experts \footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}), it is useful to be able to assess the random match probability. In other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern? Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, same models over different years, etc)

From a statistical perspective, random match probability could be calculated by sampling a large number of shoes from a given population and counting the number of shoes in the sample whose tread patterns are similar to the given print. In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area. Such a scanner could take pictures of shoes walking past, and those images could then be used to create a database of shoes relevant to the local area. \mt{Are cameras more robust/weatherproof than pressure scanners? Why images instead of EverOS impressions?} \mt{(Note: a grant proposal for such a scanner has been approved by NIJ, and there currently exists a working prototype.)} Assuming that such a database of images is available, the task becomes assessing the similarity of the shoe tread patterns in the images. This is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
